---
title: "Commercial Building Electrification at Scale"
output: 
  html_document: default 

---

```{r, echo = FALSE, message = FALSE, warning = FALSE}

###############################################################################
#                             Set up environment                                #
###############################################################################

my_packages <- c("tidyverse", "vroom" , "janitor" , "glue" ,  "tsibble" , "tidytext","lubridate", "fable", "tsibbledata", "ggplot2", "forecast", "tseries", "rio", "zoo", "readxl", "tsibbledata", "knitr", "kablExtra", "formattable") 
invisible( lapply(my_packages, require, character.only = TRUE))

#Choose one working site:
#  place <- "Home"
 place <-  "work"


#Set proper working Dir 
if (place == "Home"){setwd("C:/Users/paulr/Documents/R/cbwg")} else
{setwd("C:/Users/prode/Documents/R/cbwg")}

# Check for data directory and if one is not present then make it
if (!file.exists("data")) {
  dir.create("data")}

rm(place, my_packages )
  
```

The goal with this data project is to use the LL84 benchmarking to
identify the thermal potential in each of the building typologies. 1.
Identify the kBTU's of gas, oil and district steam used (thermal fuel)
in the commercial stock. 2. Allocate the thermal fuel among building
typologies. 3. Show ranking of typologies by sf and thermal fule volumn.
3. Show steam in manhattan as a breakout.



```{r echo=FALSE, message=FALSE, warning=FALSE}

# Read in data
L84_2017 <- read_excel("data/2017_nyc_benchmarking.xlsx", sheet = "Information and Metrics", col_names = TRUE, na = "NA"  ) %>%  rbind( read_excel("data/2017_nyc_benchmarking.xlsx", sheet = "Multi-BBL", col_names = TRUE, na = "NA"  )) %>%   
  mutate(Data = "2017") %>%
  select(1:3, 6, 8:9, 11, 14, 16:17, 25:26, 30:31, 33, 37:43, 45:46,48, 49, 51:52, 61 ) %>%
  data.frame() 


# Add Size and age bins
Building_GSF_Ranges <- c("0 to 50,000", "50,000 to 100,000", "100,000 to 500,000", "500,000 to 1,000,000", "over 1,000,000")
L84_2017$bins <- cut(L84_2017$`Self.Reported.Gross.Floor.Area..ft².`, breaks = c(0,50000, 100000, 500000, 1000000, 9000000), dig.lab = 7 , ordered_result = TRUE, labels = Building_GSF_Ranges ) 
Building_Ages <- c("pre 1942", "1942 to 1980", "1980 to 2000")
L84_2017$Year.Range <- cut(L84_2017$Year.Built, breaks = c(0, 1942, 1980, 2021),dig.lab = 4, labels = Building_Ages)
remove(Building_Ages, Building_GSF_Ranges) #clean up environment


# Add total thermal 
L84_2017 %>% rowwise() %>%  mutate("Total.Thermal (kBTU)" = sum(Fuel.Oil..1.Use..kBtu., Fuel.Oil..2.Use..kBtu., Fuel.Oil..4.Use..kBtu. , Fuel.Oil..5...6.Use..kBtu.,  Diesel..2.Use..kBtu. , Propane.Use..kBtu. , District.Steam.Use..kBtu. + District.Chilled.Water.Use..kBtu. , Natural.Gas.Use..kBtu., na.rm = TRUE) ) -> L84_2017

# Remove bad data - outliers
L84_2017_A <- L84_2017
outliers <- boxplot(L84_2017$District.Steam.Use..kBtu., plot = FALSE)$out
L84_2017_A <- L84_2017_A[-which(L84_2017_A$District.Steam.Use..kBtu. %in% outliers), ]
boxplot(L84_2017$District.Steam.Use..kBtu.)
boxplot(L84_2017_A$District.Steam.Use..kBtu.)


load( "./data/typetable.RData")
colnames(type.table) <- c("Primary.Property.Type...Self.Selected", "Property.Type")
L84_2017 <- left_join(L84_2017, type.table, by = "Primary.Property.Type...Self.Selected")
L84_2017$Property.Type <- as.factor(L84_2017$Property.Type)

L84_2017 %>%
  filter(Data != "Multi-BBL") %>%
  summarise("NGas.(kBtu)" = sum(Natural.Gas.Use..kBtu. , na.rm = TRUE ), "Oil.No1" = sum(Fuel.Oil..1.Use..kBtu., na.rm = TRUE), "Oil.No2" = sum(Fuel.Oil..2.Use..kBtu., na.rm = TRUE), "Oil,No4" = sum(Fuel.Oil..4.Use..kBtu., na.rm = TRUE), "Oil.No5&6" = sum(Fuel.Oil..5...6.Use..kBtu., na.rm = TRUE), "Diesel.(kBtu)" = sum(Diesel..2.Use..kBtu., na.rm = TRUE), "Propane.(kBtu)" = sum(Propane.Use..kBtu., na.rm = TRUE), "Steam.(kBtu)" = sum(District.Steam.Use..kBtu., na.rm = TRUE), "Elect.(kBtu)" = sum(Electricity.Use...Grid.Purchase..kBtu., na.rm = TRUE)) -> Total.Energy



```

What is the total thermal energy as reported through the Benchamrking
Law, and how does that compair to Electrical energy used in the same
buildings.

Use the information file.

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Total Energy by Fuel Type
Total.Energy %>%
  gather(key = "Fuel.Source", value = "kBtu" ) %>%
  arrange(desc(kBtu))%>%
  mutate(Fuel.Source = factor(Fuel.Source, levels = unique(Fuel.Source))) %>% 
  ggplot(aes(x = Fuel.Source, y=kBtu, fill = Fuel.Source)) +
  geom_col() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  ggtitle("Total Energy by Fuel Type") -> TotalEnergy_by_FuelType
  ggsave(filename =  "Energy_by_Type.jpeg", device = "jpeg", scale = 1, width = 6, height = 4, bg = "grey")

  TotalEnergy_by_FuelType
  
  
```





```{r, echo = FALSE, message = FALSE, warning = FALSE}

  
# Make a box plot showing fule type statics in a box plot
L84_2017 %>%
  select(Fuel.Oil..1.Use..kBtu., Fuel.Oil..2.Use..kBtu., Fuel.Oil..4.Use..kBtu., Fuel.Oil..5...6.Use..kBtu., Diesel..2.Use..kBtu., Propane.Use..kBtu., District.Steam.Use..kBtu., Natural.Gas.Use..kBtu., Electricity.Use...Grid.Purchase..kBtu.) %>%
  gather(key = "Fuel.Source", value = "kBtu") %>%
  filter(Fuel.Source == "District.Steam.Use..kBtu.") %>%
  mutate(Fuel.Source = factor(Fuel.Source, levels = unique(Fuel.Source))) %>% 
  ggplot(aes(x = Fuel.Source, y=kBtu, fill = Fuel.Source) ) +
  geom_boxplot() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
  
L84_2017 %>% select(Fuel.Oil..1.Use..kBtu., Fuel.Oil..2.Use..kBtu., Fuel.Oil..4.Use..kBtu., Fuel.Oil..5...6.Use..kBtu., Diesel..2.Use..kBtu., Propane.Use..kBtu., District.Steam.Use..kBtu., Natural.Gas.Use..kBtu., Electricity.Use...Grid.Purchase..kBtu.) %>%
  gather(key = "Fuel.Source", value = "kBtu") %>%
  filter(Fuel.Source == "District.Steam.Use..kBtu.") -> temp
summary(temp$kBtu)

  
Total.Energy%>%
  mutate("Total.Thermal.(kBtu)" = `NGas.(kBtu)`+ Oil.No1 + Oil.No2 + `Oil,No4` + `Oil.No5&6`+ `Diesel.(kBtu)`+ `Propane.(kBtu)`+ `Steam.(kBtu)`) %>%
  select(`Total.Thermal.(kBtu)`, `Elect.(kBtu)`) %>%
  gather(key = "Fuel.Source", value = "kBtu" ) %>% 
  arrange(desc(kBtu))%>%
  mutate(Fuel.Source = factor(Fuel.Source, levels = unique(Fuel.Source))) %>% 
  ggplot(aes(x = Fuel.Source, y=kBtu, fill = Fuel.Source) ) +
  geom_col() + theme(axis.text.x = element_text(angle = 60, hjust = 1))

L84_2017 %>% 
  ggplot(aes(x = `Site/Source EUI`)) +
  geom_density()

#Note how EUI ratio is lower for all electric buildings. 

L84_2017 %>%
  filter(`Total.Thermal (kBTU)`== 0) %>%
  ggplot(aes(x = `Site/Source EUI`)) +
  geom_density()

```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#EUI 
L84_2017 %>% 
  filter(`Site.EUI..kBtu.ft².` < 500) %>%
  ggplot(aes( x = Property.Type, y = EUI, color = Property.Type)) +
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

L84_2017 %>% 
  ggplot(aes(x = `EUI (Thermal/Total)`)) + 
  geom_density() +
  geom_vline(xintercept=mean(L84_2017$`EUI (Thermal/Total)` , na.rm = TRUE) ,color="blue", linetype="dashed", size=1)

#EUI by type and range
L84_2017 %>% filter(`Site.EUI..kBtu.ft².` < 500) %>%
  filter(Property.Type == "Office" | Property.Type == "Multifamily" ) %>%
  ggplot(aes( x = Year.Range, y = EUI, color = Property.Type, fill=Property.Type)) +
  geom_boxplot()

L84_2017 %>% 
  select(`Self.Reported.Gross.Floor.Area..ft².`, Number.of.Buildings, District.Steam.Use..kBtu.) %>%
  na.omit() 
#  summarise(Gross.SF = sum(Self.Reported.Gross.Floor.Area..ft².), Buildings = 
# sum(Number.of.Buildings), Steam = sum(District.Steam.Use..kBtu.))

```

